---
title: "H1 — Team accuracy (primary, directional)"
output:
  html_document:
    toc: true
---

## Locked-formula declaration

The H1 model formula is loaded from `_constants.R` (`FORMULA_H1`) and
**must** match `study/preregistration.md` §9 character-for-character.
The Python invariant test
(`backend/tests/test_pipeline_invariants.py::TestH1FormulaLocked`)
fails CI if the strings drift.

```{r setup}
suppressPackageStartupMessages({
  library(lme4)
  library(parameters)
  library(emmeans)
  library(effectsize)
  library(broom.mixed)
})

source("_constants.R")
set.seed(SEED)

data <- readRDS(file.path("data", "processed", "study.rds"))

# Confirm the locked formula matches the constant. If a contributor
# accidentally edits the literal model line below, this assertion
# tells them to update _constants.R *and* preregistration.md instead
# (which is then a registered deviation).
LITERAL_FORMULA <- "team_accuracy ~ tool * difficulty + trial_index + (1 + tool | participant) + (1 | item)"
stopifnot(identical(LITERAL_FORMULA, FORMULA_H1))
```

## Fit

```{r fit-h1}
fit_h1 <- glmer(
  formula = as.formula(FORMULA_H1),
  family = binomial(link = "logit"),
  data = data,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(fit_h1)
```

## Effect size — Westfall-Kenny-Judd 2014 d (95 % profile CI)

```{r effect-size}
# Pull the tool=hypothex_tips fixed-effect coefficient on the logit scale
# and convert to Westfall-Kenny-Judd standardised d:
#   d = b / sqrt(σ²_p + σ²_i + π²/3)
fixef_h1 <- fixef(fit_h1)
varcomp <- VarCorr(fit_h1)
sigma2_p <- attr(varcomp$participant, "stddev")[1]^2
sigma2_i <- attr(varcomp$item, "stddev")[1]^2
b_tips <- fixef_h1["toolhypothex_tips"]
d_tips <- b_tips / sqrt(sigma2_p + sigma2_i + (pi^2 / 3))

# Profile CI on the logit-scale coefficient, then converted to d
ci_b <- confint(fit_h1, parm = "toolhypothex_tips", method = "profile",
                 level = 0.95)
ci_d <- ci_b / sqrt(sigma2_p + sigma2_i + (pi^2 / 3))

cat(sprintf("H1 effect size: d = %.3f, 95%% CI [%.3f, %.3f]\n",
            d_tips, ci_d[1], ci_d[2]))
```

## p-value (used in Holm correction below)

```{r p-h1}
p_h1 <- summary(fit_h1)$coefficients["toolhypothex_tips", "Pr(>|z|)"]
cat(sprintf("H1 raw p-value: %.4g\n", p_h1))
```

## Holm correction across the H1 / H3 / H4 family

```{r holm}
# H3 and H4 p-values are produced by `03-secondary-h3-h4-glmer.Rmd`;
# here we collect them via cached output. The Holm family is locked
# in _constants.R as HOLM_FAMILY.
p_h3 <- readRDS(file.path("cache", "p_h3.rds"))
p_h4 <- readRDS(file.path("cache", "p_h4.rds"))

p_holm <- p.adjust(c(p_h1 = p_h1, p_h3 = p_h3, p_h4 = p_h4), method = "holm")
print(p_holm)
```

## Save fit + summary

```{r save}
dir.create("cache", showWarnings = FALSE)
saveRDS(fit_h1, file.path("cache", "fit_h1.rds"))
saveRDS(list(d = d_tips, ci = ci_d, p = p_h1, p_holm = p_holm),
        file.path("cache", "h1_summary.rds"))
```
