---
title: "H3 / H4 — Plausibility + cherry-picking risk (secondary)"
output:
  html_document:
    toc: true
---

## Locked formulas

H3 and H4 use the *same* fixed-effects + random-effects structure as H1
(VAL-040 §9), with the dependent variable swapped:

  - H3: `ynn5_dtw_plausibility_mean` (per VAL-003)
  - H4: `cherry_picking_risk_score` (per VAL-013)

Both are continuous, so we use `lmer` (Gaussian) rather than `glmer`.
BH q = 0.10 within the secondary family.

```{r setup}
suppressPackageStartupMessages({
  library(lme4)
  library(parameters)
  library(broom.mixed)
})

source("_constants.R")
set.seed(SEED)

data <- readRDS(file.path("data", "processed", "study.rds"))
```

## H3 — yNN_5 plausibility (per-trial outcome)

```{r fit-h3}
fit_h3 <- lmer(
  ynn5_dtw_plausibility_mean ~ tool * difficulty + trial_index +
    (1 + tool | participant) + (1 | item),
  data = data,
  control = lmerControl(optimizer = "bobyqa")
)
summary(fit_h3)

p_h3 <- summary(fit_h3)$coefficients["toolhypothex_tips", "Pr(>|t|)"]
b_h3 <- fixef(fit_h3)["toolhypothex_tips"]
cat(sprintf("H3: b = %.3f, p = %.4g\n", b_h3, p_h3))

dir.create("cache", showWarnings = FALSE)
saveRDS(p_h3, file.path("cache", "p_h3.rds"))
saveRDS(fit_h3, file.path("cache", "fit_h3.rds"))
```

## H4 — Cherry-picking risk score (session-level outcome)

The score is one number per participant (session-level metric per
VAL-013), so the model collapses the trial-level random effects.
We still keep the participant random intercept for interpretability,
but `(1 + tool | participant)` is dropped because each participant has
exactly one cherry-picking score.

```{r fit-h4}
sess <- data %>%
  dplyr::group_by(participant, tool, item, trial_index) %>%
  dplyr::summarise(
    cherry_picking_risk_score = first(cherry_picking_risk_score),
    .groups = "drop"
  )

fit_h4 <- lmer(
  cherry_picking_risk_score ~ tool * difficulty + (1 | participant),
  data = sess,
  control = lmerControl(optimizer = "bobyqa")
)
summary(fit_h4)

p_h4 <- summary(fit_h4)$coefficients["toolhypothex_tips", "Pr(>|t|)"]
b_h4 <- fixef(fit_h4)["toolhypothex_tips"]
cat(sprintf("H4: b = %.3f, p = %.4g\n", b_h4, p_h4))

saveRDS(p_h4, file.path("cache", "p_h4.rds"))
saveRDS(fit_h4, file.path("cache", "fit_h4.rds"))
```

## BH correction across the secondary family (H3, H4)

```{r bh}
p_secondary <- c(p_h3 = p_h3, p_h4 = p_h4)
p_bh <- p.adjust(p_secondary, method = "BH")
print(p_bh)
cat(sprintf("BH q threshold = %.2f\n", BH_Q))
saveRDS(list(p = p_secondary, p_bh = p_bh, q = BH_Q),
        file.path("cache", "secondary_bh.rds"))
```
