---
title: "Bayesian companion — informational, not gating"
output:
  html_document:
    toc: true
---

Companion `brms` model with the same fixed-effects structure as H1.
Per VAL-040 §10:
  - `Normal(0, 1)` priors on standardised effects.
  - Posterior probability that the effect exceeds **ROPE = ± 0.1 d**
    (Kruschke 2013).
  - **Robustness check only** — never used to overturn or confirm the
    frequentist H1 outcome.

AC reporting requirements (line 159):
  - Posterior mean.
  - 95 % credible interval.
  - Posterior probability outside ROPE.
  - R-hat ≤ 1.01 for all parameters.
  - ESS ≥ 1000 for all bulk + tail.

```{r setup}
suppressPackageStartupMessages({
  library(brms)
  library(posterior)
  library(bayestestR)
})

source("_constants.R")
set.seed(SEED)

data <- readRDS(file.path("data", "processed", "study.rds"))
```

## Priors

```{r priors}
priors <- c(
  prior(normal(0, 1),       class = "b"),         # standardised effects
  prior(student_t(3, 0, 2.5), class = "Intercept"),
  prior(student_t(3, 0, 2.5), class = "sd"),
  prior(lkj(2),              class = "cor")
)
```

## Fit

```{r fit-bayes}
fit_bayes <- brm(
  formula = as.formula(FORMULA_H1),
  family = bernoulli(),
  prior = priors,
  data = data,
  chains = BRMS_CHAINS,
  iter = BRMS_ITER,
  warmup = BRMS_WARMUP,
  control = list(adapt_delta = BRMS_ADAPT_DELTA),
  seed = SEED,
  file = file.path("cache", "fit_bayes.rds")
)
```

## Convergence diagnostics (LOCKED thresholds)

```{r diagnostics}
ds <- summary(fit_bayes)$fixed
rhat_max <- max(ds$Rhat, na.rm = TRUE)
ess_bulk_min <- min(ds$Bulk_ESS, na.rm = TRUE)
ess_tail_min <- min(ds$Tail_ESS, na.rm = TRUE)

cat(sprintf("R-hat max = %.4f (threshold ≤ %.2f) — %s\n",
            rhat_max, BRMS_R_HAT_MAX,
            if (rhat_max <= BRMS_R_HAT_MAX) "PASS" else "FAIL"))
cat(sprintf("Bulk ESS min = %d (threshold ≥ %d) — %s\n",
            as.integer(ess_bulk_min), BRMS_ESS_MIN,
            if (ess_bulk_min >= BRMS_ESS_MIN) "PASS" else "FAIL"))
cat(sprintf("Tail ESS min = %d (threshold ≥ %d) — %s\n",
            as.integer(ess_tail_min), BRMS_ESS_MIN,
            if (ess_tail_min >= BRMS_ESS_MIN) "PASS" else "FAIL"))

if (rhat_max > BRMS_R_HAT_MAX || min(ess_bulk_min, ess_tail_min) < BRMS_ESS_MIN) {
  warning(
    "Bayesian companion convergence diagnostics did not meet pre-",
    "registered thresholds — log a deviation in study/deviations.md ",
    "and report in the paper."
  )
}
```

## Posterior summary + ROPE

```{r posterior}
post_b <- posterior::as_draws_df(fit_bayes)$b_toolhypothex_tips
post_mean <- mean(post_b)
post_ci <- quantile(post_b, c(0.025, 0.975))

# Convert from logit-scale b to standardised d for the ROPE comparison.
varcomp <- VarCorr(fit_bayes)
sigma2_p <- attr(varcomp$participant, "stddev")[1]^2
sigma2_i <- attr(varcomp$item, "stddev")[1]^2
denom <- sqrt(sigma2_p + sigma2_i + (pi^2 / 3))
post_d <- post_b / denom
prob_outside_rope <- mean(abs(post_d) > ROPE_D)

cat(sprintf(
  "Posterior mean (b) = %.3f, 95%% CI [%.3f, %.3f]\n",
  post_mean, post_ci[1], post_ci[2]
))
cat(sprintf(
  "Posterior d mean = %.3f, Pr(|d| > %.2f) = %.3f\n",
  mean(post_d), ROPE_D, prob_outside_rope
))

dir.create("cache", showWarnings = FALSE)
saveRDS(list(
  post_mean = post_mean,
  post_ci = post_ci,
  post_d_mean = mean(post_d),
  prob_outside_rope = prob_outside_rope,
  rhat_max = rhat_max,
  ess_bulk_min = ess_bulk_min,
  ess_tail_min = ess_tail_min,
  rope_d = ROPE_D
), file.path("cache", "bayes_summary.rds"))
```
