---
title: "Exploratory analyses (declared in VAL-040 §12)"
output:
  html_document:
    toc: true
---

> **Reminder.** These contrasts are *exploratory* in label; all are
> declared up-front in VAL-040 §12. No NHST claim; report estimates +
> 95 % CIs only. New exploratory contrasts added here would be a
> registered deviation.

```{r setup}
suppressPackageStartupMessages({
  library(dplyr)
  library(lme4)
  library(parameters)
})

source("_constants.R")
set.seed(SEED)

data <- readRDS(file.path("data", "processed", "study.rds"))
```

## H5 — tip engine increases shape coverage

```{r h5}
fit_h5 <- lmer(
  shape_coverage_fraction ~ tool + (1 | participant),
  data = data %>%
    dplyr::group_by(participant, tool) %>%
    dplyr::summarise(
      shape_coverage_fraction = first(shape_coverage_fraction),
      .groups = "drop"
    )
)
print(parameters::model_parameters(fit_h5, ci = 0.95))

dir.create("cache", showWarnings = FALSE)
saveRDS(fit_h5, file.path("cache", "fit_h5.rds"))
```

## Correlation: cherry-picking risk vs. trust calibration

```{r corr-cp-trust}
session_level <- data %>%
  dplyr::group_by(participant, tool) %>%
  dplyr::summarise(
    cherry_picking_risk_score = first(cherry_picking_risk_score),
    trust_calibration_brier = first(trust_calibration_brier),
    .groups = "drop"
  )

cor_test <- cor.test(
  session_level$cherry_picking_risk_score,
  session_level$trust_calibration_brier,
  method = "pearson"
)
print(cor_test)
saveRDS(cor_test, file.path("cache", "exploratory_cp_trust_corr.rds"))
```

## Per-shape-primitive accuracy

```{r per-shape}
# `shape_primitive` per item is in items.json; data joins it onto trials.
per_shape <- data %>%
  dplyr::group_by(tool, shape_primitive) %>%
  dplyr::summarise(
    accuracy = mean(team_accuracy),
    n = dplyr::n(),
    .groups = "drop"
  )
print(per_shape)
saveRDS(per_shape, file.path("cache", "exploratory_per_shape.rds"))
```

## Tip-modality effect within `hypothex_tips` only

```{r tip-modality}
tips_only <- data %>%
  dplyr::filter(tool == "hypothex_tips" & !is.na(tip_modality))

if (nrow(tips_only) >= 30) {
  fit_modality <- glmer(
    team_accuracy ~ tip_modality + (1 | participant) + (1 | item),
    family = binomial(),
    data = tips_only
  )
  print(parameters::model_parameters(fit_modality, ci = 0.95))
  saveRDS(fit_modality, file.path("cache", "exploratory_tip_modality.rds"))
} else {
  cat("Insufficient hypothex_tips trials with tip_modality recorded;",
      "skipping per-modality contrast.\n")
}
```

## Domain-pack effect on each primary outcome

```{r domain}
for (outcome in PRIMARY_OUTCOMES) {
  cat("---", outcome, "---\n")
  fm <- as.formula(paste0(outcome, " ~ tool * domain_pack + (1 | participant) + (1 | item)"))
  fit_dom <- tryCatch(
    if (outcome == "team_accuracy") {
      glmer(fm, data = data, family = binomial())
    } else {
      lmer(fm, data = data)
    },
    error = function(e) { cat("  fit failed:", conditionMessage(e), "\n"); NULL }
  )
  if (!is.null(fit_dom)) {
    print(parameters::model_parameters(fit_dom, ci = 0.95))
    saveRDS(fit_dom, file.path("cache", paste0("exploratory_domain_", outcome, ".rds")))
  }
}
```
