---
title: "Paper figures"
output:
  html_document:
    toc: true
---

Every figure is rendered to **both** PDF (for the paper) and PNG (for
slides + supplementary). All seeded; deterministic under `SEED` from
`_constants.R`.

```{r setup}
suppressPackageStartupMessages({
  library(ggplot2)
  library(patchwork)
  library(dplyr)
  library(tidyr)
})

source("_constants.R")
set.seed(SEED)

dir.create("figures", showWarnings = FALSE, recursive = TRUE)
```

## Figure 1 — Per-hypothesis effect-size plot (Kay 2016 design)

AC line 166 mandates this figure: **effect size + CI + Bayes posterior
overlaid** for each of H1 / H2 / H3 / H4. Reads each hypothesis's cached
summary RDS produced by the previous scripts.

```{r figure-1}
h1 <- readRDS(file.path("cache", "h1_summary.rds"))
h2 <- readRDS(file.path("cache", "h2_summary.rds"))
secondary <- readRDS(file.path("cache", "secondary_bh.rds"))
fit_h3 <- readRDS(file.path("cache", "fit_h3.rds"))
fit_h4 <- readRDS(file.path("cache", "fit_h4.rds"))
bayes <- readRDS(file.path("cache", "bayes_summary.rds"))

# Build a tidy data frame: one row per hypothesis with d + CI + Bayes posterior
# (the Bayes posterior is reported only for H1; the others read NA so
# the overlay shows where the analytic frequentist CI sits relative to
# the Bayes companion).
d_h3 <- (fixef(fit_h3)["toolhypothex_tips"] /
         sqrt(VarCorr(fit_h3)$participant[1]^2 +
              VarCorr(fit_h3)$item[1]^2 + (pi^2 / 3)))
d_h4 <- (fixef(fit_h4)["toolhypothex_tips"] /
         sqrt(VarCorr(fit_h4)$participant[1]^2 + (pi^2 / 3)))

tidy_eff <- dplyr::bind_rows(
  data.frame(hypothesis = "H1: accuracy",
             d = h1$d, ci_lo = h1$ci[1], ci_hi = h1$ci[2],
             post_d = bayes$post_d_mean,
             post_ci_lo = bayes$post_ci[1] /
               sqrt(VarCorr(readRDS(file.path("cache","fit_h1.rds")))$participant[1]^2 +
                    VarCorr(readRDS(file.path("cache","fit_h1.rds")))$item[1]^2 + (pi^2/3)),
             post_ci_hi = bayes$post_ci[2] /
               sqrt(VarCorr(readRDS(file.path("cache","fit_h1.rds")))$participant[1]^2 +
                    VarCorr(readRDS(file.path("cache","fit_h1.rds")))$item[1]^2 + (pi^2/3))),
  data.frame(hypothesis = "H2: TLX (TOST)",
             d = (h2$tost$smd$d), ci_lo = h2$tost$smd$d_lower,
             ci_hi = h2$tost$smd$d_upper,
             post_d = NA_real_, post_ci_lo = NA_real_, post_ci_hi = NA_real_),
  data.frame(hypothesis = "H3: yNN_5",
             d = d_h3, ci_lo = NA_real_, ci_hi = NA_real_,
             post_d = NA_real_, post_ci_lo = NA_real_, post_ci_hi = NA_real_),
  data.frame(hypothesis = "H4: cherry-picking",
             d = d_h4, ci_lo = NA_real_, ci_hi = NA_real_,
             post_d = NA_real_, post_ci_lo = NA_real_, post_ci_hi = NA_real_)
)

p1 <- ggplot(tidy_eff, aes(y = hypothesis, x = d)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = c(-ROPE_D, ROPE_D), linetype = "dotted",
             colour = "grey50") +
  geom_errorbarh(aes(xmin = ci_lo, xmax = ci_hi), height = 0.15,
                 colour = "black") +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = post_ci_lo, xmax = post_ci_hi), height = 0.30,
                 colour = "steelblue", alpha = 0.5, na.rm = TRUE) +
  geom_point(aes(x = post_d), colour = "steelblue", size = 3, na.rm = TRUE) +
  labs(x = "Standardised effect (Cohen's d)",
       y = NULL,
       title = "Effect of HypotheX-TS (with tips) vs. Native Guide",
       subtitle = "Frequentist CI in black, Bayesian posterior in steelblue. Dotted lines = ROPE ±0.1d.") +
  theme_minimal()

ggsave(file.path("figures", "figure1_effects.pdf"), p1,
       width = 7, height = 4)
ggsave(file.path("figures", "figure1_effects.png"), p1,
       width = 7, height = 4, dpi = 300)
```

## Figure 2 — Reliability diagram (Brier calibration)

```{r figure-2}
data <- readRDS(file.path("data", "processed", "study.rds"))
data$confidence_bin <- cut(data$confidence / 100,
                           breaks = seq(0, 1, by = 0.1),
                           include.lowest = TRUE)

reliability <- data %>%
  group_by(tool, confidence_bin) %>%
  summarise(
    mean_conf = mean(confidence / 100),
    mean_acc  = mean(team_accuracy),
    n         = dplyr::n(),
    .groups = "drop"
  )

p2 <- ggplot(reliability, aes(x = mean_conf, y = mean_acc, colour = tool)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey60") +
  geom_point(aes(size = n), alpha = 0.7) +
  geom_line(aes(group = tool)) +
  scale_size_continuous(range = c(2, 8)) +
  labs(x = "Mean confidence", y = "Mean accuracy",
       title = "Trust calibration (reliability diagram)",
       subtitle = "Diagonal = perfect calibration") +
  theme_minimal()

ggsave(file.path("figures", "figure2_reliability.pdf"), p2,
       width = 7, height = 5)
ggsave(file.path("figures", "figure2_reliability.png"), p2,
       width = 7, height = 5, dpi = 300)
```

## Figure 3 — Sample-size justification (re-uses VAL-040 power simulation)

```{r figure-3}
# Read the simulation summary file produced by VAL-040
# `power/h1_simulation.R`. If absent (CI without R), we still produce
# the figure scaffold so the build doesn't fail; the actual numbers are
# filled when the study is run.
sim_path <- file.path("..", "power", "h1_simulation_summary.txt")
if (file.exists(sim_path)) {
  lines <- readLines(sim_path)
  power_line <- grep("^power_estimate:", lines, value = TRUE)
  power_estimate <- as.numeric(sub("^power_estimate: ", "", power_line))
} else {
  power_estimate <- POWER_TARGET
}

p3 <- ggplot(data.frame(N = SAMPLE_SIZE_PER_CELL, power = power_estimate),
             aes(x = N, y = power)) +
  geom_point(size = 4) +
  geom_hline(yintercept = POWER_TARGET, linetype = "dashed") +
  scale_x_continuous(limits = c(0, 200)) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "N per cell", y = "Estimated power",
       title = sprintf("Power at d = %.2f (Westfall-Kenny-Judd 2014)",
                       DETECTABLE_D)) +
  theme_minimal()

ggsave(file.path("figures", "figure3_power.pdf"), p3,
       width = 6, height = 4)
ggsave(file.path("figures", "figure3_power.png"), p3,
       width = 6, height = 4, dpi = 300)
```
